Hyper-linked Communications: WebRTC enabled asynchronous collaboration

### 2016-05-13 tese2/conclusions.tex
[ADD] final version of statistics

\section{Summary}
\label{section:summary}
In this thesis, our goal was the development of a web application using \emph{WebRTC} that complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time. Another important goal of this project is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.

Having this in mind, we started to analyze the technologies necessary to implement our solution. In a first glance, we have analyzed the problems that real time communication applications are facing when using the current Internet infrastructure. Next we have we have analyzed the main components of the \emph{WebRTC} stack. Furthermore, we have analyzed how we could implement the signalization component that \emph{WebRTC} does not not define. Hence, with the transport component being analyzed, we have studied how we could expose content in a synchronized and interactive way. Moreover, we have performed an analysis on the kinds of media types the particularities of each one when manipulating time. Lastly, on the state of the art context, we have analyzed existing frameworks and libraries that could help us implementing a collaborative environment.

@@

Although we have used our own signaling protocol on our implementation, the first steps we have made during the development of our prototype were using \ac{XMPP}, we have discovered the drawbacks of this approach in practice and we moved our learnings to the state of the art. Besides starting our prototype using this approach, we have used \emph{strophe.js} for performing communications with \ac{XMPP} servers. We have found and fixed a bug present on \emph{strophe.js} library that silenced an error raised by duplicated user registrations.

As a consequence of not using \ac{XMPP} for the signaling protocol, we have defined our own protocol based on \emph{WebSockets} for performing the stream between clients and \ac{KMS}. From the \ac{KMS} functionalities, our solution uses the recording feature, composite endpoints that mix multiple audio and video streams into a single one and \emph{QR code} detection for creating hyper content.

@@

\section{Future Work}
\label{section:future}
	Playing back video with a faster rate is not possible using the current version of \ac{KMS} which we expect the availability of that feature in a near future but we have proposed an alternative way to implement faster playback by using \emph{ffmpeg} to convert the video before playing it.

	Although we have tested our solution in a powerful machine for the current time, our performance tests revealed that the streaming component uses a lot of resources. We left for a future work a deep analysis on the scalability of our system which we have proposed different approaches but we have not tested them.

>>>>

\section{Summary}
\label{section:summary}
In this thesis, our goal was the development of a web application using \emph{WebRTC} that complements current audio, text and video communications, in order to create rich and collaborative interfaces with the ability to add more content on a future time. Another important goal of this project is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.

Having this in mind, we started by analyzing the technologies necessary to implement our solution. In a first glance, we have analyzed the problems that real time communication applications face when using the current Internet infrastructure. Next, we analyzed the main components of the \emph{WebRTC} stack. Furthermore, we analyzed how we could implement the signaling component that \emph{WebRTC} does not not define. Hence, with the transport component analyzed, we have studied how we could expose content in a synchronized and interactive way. Moreover, we have performed an analysis on the kinds of media types ant the particularities of each one when manipulating time. Lastly, on the state of the art context, we analyzed existing frameworks and libraries that could help us implementing a collaborative environment.

@@

Although we have used our own signaling protocol we have made our first steps using \ac{XMPP} and we have discovered the drawbacks of this approach. Dor performing communications with \ac{XMPP} servers, we used \emph{strophe.js} which we found to have a bug that we have fixed.

@@

\section{Future Work}
\label{section:future}
	Playing back video with a faster rate is not possible using the current version of \ac{KMS}. Even though we expect the availability of that feature in a near future but we have proposed an alternative way to implement faster playback by using \emph{ffmpeg} to convert the video before playing it.

	Although we have tested our solution in a powerful machine for the current time, our performance tests revealed that the streaming component uses a lot of resources. We left for a future work a deep analysis on the scalability of our system for which we have proposed different approaches but we have not tested them.

### 2016-05-13 tese2/evaluation.tex
[ADD] final version of statistics

   Figure \ref{fig:test_hour_mem} shows the memory used by our solution with seven clients simultaneously on the same conference room during one hour. As we can observe \ac{KMS} memory usage stabilizes.

@@

      In order to the usability of our solution, we haver performed usability tests with the help of real users with different backgrounds and ages.

      We handled a guide to the users with five tasks to perform, the metrics we used for each task were: number of clicks, number of errors including a description and time spent.

      The tasks we asked to the user are:

      \begin{enumerate}
      \item Login into the system with the provided credentials and accept the received friendship request. Find the coordinator's brother and add him as a friend.

@@

      \item Enter in the conference room correspondent to the third task as an observer and navigate to the specified annotation. Watch the video until a list of topics is overlayed in the video and choose one of them and leave the conference room.

      \item Enter in the conference room correspondent to the fourth task by sharing your camera and create a time annotation in the instant of time when you entered, navigate to the current time, search for the annotation and leave the conference room.

      \item Enter in the conference room correspondent to the fourth task by sharing your camera and add a subtitle in the video, preview it, specify an interval of time and save it. Then show the provided \emph{QR code} to the camera and leave the leave the conference room.
      \end{enumerate}

      The goal of the first task consists on evaluating the interface for authentication and friendship management.

      The second task is used to make the users familiar with our tools for collaborative content edition.

@@

      The goal of the fifth task is to create synchronized overlayed content, demonstrate the difficulties of such task and present an easier way to synchronize content.

@@

In this section we present the results of our usability tests. The first time we tested our solution by providing tasks to users, we have observed that our solution was not perfect. We have faced usability problems during our tests, he had to improve our solution's usability and start the tests again.

@@

In the first testing phase we have performed just three tests with users and ceased for improvements.

Both users that tested our solution had difficulties on the first task when finding new friendship requests and adding new friends, as friendship request were possible to access only from the top bar and new user profiles were only accessible through the search bar. In order to solve this problem we have added a menu for adding new users in the friends list and another near the friends list as it can be seen on figure \ref{fig:test_ui_01_02_03}.

@@

On the second task one user clicked on the friends list for adding a friend to room and by consequence made him leave the conference room to the user profile. We solved this problem by showing a pop-up with the user profile and an additional button to add the user to the conference room as it can be seen on figure \ref{fig:test_ui_04}.

@@

On the third task we noticed that both users were not expecting to search the content directly from the search bar, instead they used the time-line.

On the fourth task both users didn't placed their annotations on right place, we only provide one way to place annotations, which is by dragging them on the time-line. 
After creating an annotation users also didn't understand that they had to save their changes on the time-line, we solved this problem by saving automatically each time the user changed any annotation, the old user interface can be seen on figure \ref{fig:test_ui_05}.

@@

Both users had success with fifth task but they revealed difficulties choosing the time interval for the hyper content, we are aware that this is a difficult task and that's why we have introduced the \emph{QR code} content creation mechanism which was very easy to learn.

@@

\begin{itemize}
\item Giving feedback to user when saving content and collaborative editors.
\item Support file upload on our chat.
\item Starting with our time-line more zoomed in order to give more sensation of time.
\item Create content with starting time and duration instead ending time.
\end{itemize}

@@

On the second phase of our user interface tests, in a general view, we have noticed great improvements on the learning time.

@@

From the data that we have collected, namely the task duration (figure \ref{fig:user_times}), number of clicks (figure \ref{fig:user_clicks}), number of errors (figure \ref{fig:user_errors}), task difficulty (figure \ref{fig:user_diffs}) and qualitative evaluation (figure \ref{fig:user_evals}), we have calculated the confidence intervals in order to understand the most plausible values for each metric.

@@

Where $\alpha=0.05$, $n$ is the number of samples (which may not coincide with the number of tests due to failed tasks), $\bar{x}$ is the sample mean, $x_i$ is the sample value and $s$ is the sample standard deviation that is given by:
\begin{equation}
s=\sqrt{\frac{1}{n-1}\Sigma^{n}_{i=1}(x_i-\bar{x})^2}
\end{equation}

@@

Accordingly to figure \ref{fig:user_evals}, we can observe that most users had less difficulties with the first two tasks, which represents types of tasks that most users are familiar with. As soon as users had to navigate in time, manipulate annotations and create content (respectively \emph{task3}, \emph{task4} and \emph{task5}) we have observed that they revealed more difficulties. Most of those difficulties, based on the users feedback, were mainly due to those concepts not being familiar to them.

@@

Although we had relatively bad results with some users, we have explained to the users that could not conclude the tasks or performed incorrectly the most efficient way to perform the requested tasks. Some users suggested to display more hints in order to achieve a faster learning but after all they were impressed and gave us a better evaluation (as seen on figure \ref{fig:user_evals}).

@@

Most users gave us worse evaluations on our user interface layout and content editor, which was due to having a lot of tools present in the same web page and some of them were hidden due the screen size. In some cases users had to scroll down in order to find the tools they were looking for.

Another weak aspect was our content editor, which in fact we recognize the difficulty to handle with, most due to the amount of information that is necessary to create a synchronized content (starting time, duration and content itself). Some users have suggested that the content should also be present on the time-line so they could be easily dragged and resized (on time).

We are aware that placing content on the time-line will reduce our solutions performance, especially when there is a relatively big amount of content due to the content that is present on the time-line being loaded all in once. Although, we recognize that for some cases (relatively low amount of content), displaying the content on the time-line could not have a great impact on our solutions performance.

In conclusion, 100\% of our testers though that our solution was an innovation and 92.31\% recommended using our solution.

@@

The performance tests revealed unexpected results which were crucial to find and solve performance issues. Another important aspect of our performance test results, were the understanding of how data flows between our system modules.

@@

In order to understand how good the results were for each performed task, we have defined optimal values for the different metrics. In general the results reveal a significant difference between observed values and optimal values which can be explained by the users lack of experience with systems such as we have created.

>>>>

   Figure \ref{fig:test_hour_mem} shows the memory used by our solution with seven clients simultaneously on the same conference room during one hour. As we can observe, \ac{KMS} memory usage stabilizes.

@@

      In order to evaluate the usability of our solution, we have performed usability tests with the help of real users with different backgrounds and ages.

      We handed a guide to the users with five tasks to perform. The metrics we used for each task were: number of clicks, number of errors (including a description) and time spent.

      The tasks we asked the user to perform were:

      \begin{enumerate}
      \item Login into the system with the provided credentials and accept the received friendship request. Find a given user by its name and add him as a friend.

@@

      \item Enter a conference room correspondent as an observer and navigate to the specified annotation. Watch the video until a list of topics is overlayed in the video and choose one of them and leave the conference room.

      \item Enter a conference room by sharing their camera and create a time annotation in the instant of time when they entered, navigate to the current time, search for the annotation and leave the conference room.

      \item Enter a conference room by sharing their camera and add a subtitle in the video, preview it, specify an interval of time and save it. Then show the provided \emph{QR code} to the camera and leave the leave the conference room.
      \end{enumerate}

      The goal of the first task is to evaluating the interface for authentication and friendship management.

      The second task is used to evaluate our tools for collaborative content edition.

@@

      The goal of the fifth task is to create synchronized overlayed content and present an easier way to synchronize content.

@@

In this section we present the results of our usability tests. The first time we tested our solution by providing tasks to users, we have observed that our solution was not perfect. Having faced usability problems during our tests, he had to improve our solution's usability and start the tests again.

@@

In the first testing phase we have performed tests with just three users and ceased for improvements.

The initial users that tested our solution had difficulties in the first task when finding new friendship requests and adding new friends, as friendship request were only possible to access from the top bar and new user profiles were only accessible through the search bar. In order to solve this problem we have added a menu for adding new users in the friends list and another near the friends list as it can be seen on figure \ref{fig:test_ui_01_02_03}.

@@

On the second task one user clicked on the friends list for adding a friend to a room and, as a consequence, left the conference room to the user profile. We solved this problem by showing a pop-up with the user profile and an additional button to add the user to the conference room as it can be seen on figure \ref{fig:test_ui_04}.

@@

On the third task we noticed that users were not expecting to search the content directly from the search bar, instead they used the time-line.

On the fourth task the users didn't placed their annotations on right place. We only provide one way to place annotations, which is by dragging them on the time-line. 
After creating an annotation, users also didn't understand that they had to save their changes on the time-line. We solved this problem by saving automatically each time the user changed any annotation. The old user interface can be seen on figure \ref{fig:test_ui_05}.

@@

The three users had success with the fifth task, but they revealed difficulties choosing the time interval for the hyper content. We are aware that this is a difficult task and that's why we have introduced the \emph{QR code} content creation mechanism which was very easy to learn (with a \emph{QR code} already printed into paper).

Although content creation through \emph{QR codes} revealed much easier to perform, this approach implies showing and, as a consequence, recording video with the \emph{QR codes}, making them visible to other users. In order to solve this problem, we could add a new video sharing mode that allow users to send video to \ac{KMS} without composing, recording and showing to other users.

@@

\begin{itemize}
\item Giving feedback to user when performing operations such as create superimposed content and save the collaborative editor content.
\item Starting with our time-line more zoomed in order to give more sensation of time passing.
\item Create content with starting time and duration, instead of ending time.
\end{itemize}

@@

On the second phase of our user interface tests, in a general way, we have noticed great improvements on the learning time.

@@

From the data that we have collected, namely the task duration (Figure \ref{fig:user_times}), number of clicks (Figure \ref{fig:user_clicks}), number of errors (Figure \ref{fig:user_errors}), task difficulty (Figure \ref{fig:user_diffs}) and qualitative evaluation (Figure \ref{fig:user_evals}), we have calculated the confidence intervals in order to understand the most plausible values for each metric.

@@

Where $\alpha=0.05$ \footnote{95\% ($\alpha=0.05=1-0.95$) is a common value, used by researchers, to calculate confidence intervals}, $n$ is the number of samples (which may not coincide with the number of tests due to failed tasks), $\bar{x}$ is the sample mean, $x_i$ is the sample value and $s$ is the sample standard deviation that is given by:
\begin{equation}
s=\sqrt{\frac{1}{n-1}\Sigma^{n}_{i=1}(x_i-\bar{x})^2}
\end{equation}

@@

According to Figure \ref{fig:user_diffs}, we can observe that most users had less difficulties with the first two tasks, which represents types of tasks that most users are familiar with. As soon as users had to navigate in time, manipulate annotations and create content (respectively \emph{task3}, \emph{task4} and \emph{task5}) we observed that they revealed more difficulties. Most of those difficulties, based on the users feedback, were mainly due to those concepts not being familiar to them.

@@

As we had relatively bad results with some users, we explained to those users that could not conclude the tasks or performed them incorrectly, the most efficient way to perform the requested tasks. Some users suggested to display more hints in order to achieve a faster learning, but afterwards all were impressed and gave us a better evaluation (as seen on Figure \ref{fig:user_evals}).

@@

Most users gave us worse evaluations on our user interface layout and content editor, which was due to having a lot of tools present in the same web page and some of them being hidden due the screen size. In some cases users had to scroll down in order to find the tools they were looking for.

Another weak aspect was our content editor, which, in fact, we recognize is difficult to work with, mostly due to the amount of information that is necessary to create a synchronized content (starting time, duration and content itself). Some users have suggested that the content should also be present on the time-line so they could be easily dragged and resized (on time).

We are aware that placing content on the time-line will reduce our solutions performance, especially when there is a relatively large amount of content, due to the content that is present on the time-line being loaded all at once. Although, we recognize that for some cases (relatively low amount of content), displaying the content on the time-line could not have a great impact on our solutions performance, we choose not to implement this.

In conclusion, 100\% of our testers though that our solution was an innovation and 95\% recommended using our solution.

@@

The performance tests revealed unexpected results which were crucial to find and solve performance issues. Another important aspect of our performance test results was the understanding of how data flows between our system modules.

@@

In order to understand how good the results were for each performed task, we have defined optimal values for the different metrics. In general the results reveal a significant difference between observed values and optimal values, which can be explained by the user's lack of experience with systems such as the one we have created.

### 2016-05-13 tese2/implementation.tex
until 1h test case
[ADD] final version of statistics

\input{13_implementation_model.tex}
\input{14_implementation_signaling.tex}
\input{15_implementation_recording.tex}
\input{16_implementation_content.tex}
\input{17_implementation_deploy.tex}
\input{18_implementation_summary.tex}

>>>>

\input{13_implementation_model.tex}

\section{XMPP experiments}
Although we have used our own signaling protocol on our implementation, the first steps we made during the development of our prototype were using \ac{XMPP}. We discovered the drawbacks of this approach in practice and described our learnings in the state of the art. We have found and fixed a bug present on \emph{strophe.js} library that silenced an error raised by duplicated user registrations.

As a consequence of not using \ac{XMPP} for the signaling protocol, we have defined our own protocol based on \emph{WebSockets} for performing the stream between clients and \ac{KMS}. From the \ac{KMS} functionalities, our solution uses the recording feature, composite endpoints that mix multiple audio and video streams into a single one and \emph{QR code} detection for creating hyper content.

\input{14_implementation_signaling.tex}
\input{15_implementation_recording.tex}
\input{16_implementation_content.tex}
\input{18_implementation_summary.tex}
